Species motif participation provides unique information about species risk of extinction

### 2023-05-03 manuscript/reorder_main_text.tex
Updates from Overleaf

        To understand the relationships between motif participation and consumers' probabilities of persistence, it is helpful to understand how persistence varies with other aspects of network structure.
        For example, consumers with higher STL were less likely to persist (Fig.~\ref{fig:motifs_vs_TL_and_deg}, Table S3, \emph{Appendix S6}).
        The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.

@@

        Both the motif profile of a network and species' motif participation were related to other measures of network structure.
        Most simply, the proportion of the omnivory motif in the network's motif role increased with increasing connectance (Fig. S6, Table S5; \emph{Appendix S9}).
        The proportions of other motifs did not vary significantly with connectance and none were significantly related to network size.

        A species' motif participation varied with network size, connectance, and their interaction. 
        The proportion of the omnivory motif generally increased with increasing connectance and network size, but decreased with increasing network size in the most-connected networks (Fig. S2 and Table S2, \emph{Appendix S5}).
        The proportion of apparent competition decreased strongly with increasing connectance and more weakly (but still significantly) with increasing network size.
        The proportion of the three-species chain showed similar trends to apparent competition, while the proportion of direct competition showed very little change over the range of network size and connectance we consider (though the effects of connectance and the interaction between connectance and network size were significant).

        Species' motif participation also varied with their in-degree and trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
        Species with higher in-degree tended to have higher frequencies of participation in the omnivory motif and lower proportions of the other three motifs.
        Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains and lower proportions of omnivory and direct competition. 
        A species' persistence varied with both species in-degree and trophic level, where the most clear result is decreased persistence with increased trophic level (Fig.~\ref{fig:motifs_vs_TL_and_deg}; \emph{Appendix S6}).

        % \subsubsection*{Consistency in effects of motif participation on species persistence}

        %     When fitting regressions of persistence against motif participation for each network separately, networks with high connectance generally showed more consistent relationships between persistence and motif participation, i.e., sharper peaks of the density distributions (Fig.~\ref{fig:density_prop_C}).
        %     However, this trend varied at different levels of disturbance, especially for the direct competition and omnivory motifs.
        %     At low and medium connectances there was little change in the distribution of slopes for direct competition across all levels of disturbance. 
        %     At high connectances, however, the proportion of positive slopes was both lower overall and decreased with increasing disturbance. 
        %     For the omnivory motif these trends were reversed. 
        %     Moreover, the omnivory motif has a particularly broad distribution of slopes (i.e., an inconsistent relationship to persistence) when network connectance was low. 
        %     Network size had a much smaller effect than connectance (Fig. S4; \emph{Appendix S7}).

@@

    \begin{figure}[ht!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
        \caption{
        \textbf{A} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
        \textbf{B,E} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance. There was no significant relationship between mean persistence and network size.
        \textbf{C} Mean persistence of consumers decreased with increasing proportions of omnivory in the network's motif profile and increased with increasing proportions of other motifs. Persistence also decreased significantly with increasing probability of basal species extinction but there was no significant interaction; we therefore show trends only for $\pi$=0.1.
        \textbf{D} Persistence of consumers decreased with increasing STL at all levels of disturbance.}
        \label{fig:lm_CS}
    \end{figure}

>>>>

        To understand the relationships between motif participation and consumers' probabilities of persistence, it is helpful to understand how persistence varies with other aspects of network structure.
        Consumers with higher STL were less likely to persist (Fig.~\ref{fig:motifs_vs_TL_and_deg}, Table S3, \emph{Appendix S6}).
        The relationship between in-degree and persistence depended on the strength of disturbance: higher in-degree was associated with greater persistence when basal species had low probability of extinction but lower persistence when basal species were highly likely to go extinct.

@@

        A species' motif participation varied with its position in the network and with overall network structure (Fig.~\ref{fig:motifs_vs_TL_and_deg}).
        Species with higher in-degree tended to have higher frequencies of participation in the omnivory motif and lower proportions of the other three motifs.
        Species with higher trophic levels tended to have higher proportions of apparent competition and three-species chains and lower proportions of omnivory and direct competition.

        The frequency of participation in the omnivory motif generally increased with increasing connectance and network size, but decreased with increasing network size in the most-connected networks (Fig. S2 and Table S2, \emph{Appendix S5}).
        The frequencies of participation in the apparent competition and three-species chain motifs decreased strongly with increasing connectance but showed little change with changing network size.
        The frequency of participation in the direct competition motif decreased with increasing connectance in small networks and increased with increasing network size in highly-connected networks, but showed little change with connectance in large networks or with network size in less-connected networks.

        The network's motif profile also varied with network connectance.
        Specifically, the proportion of the profile made up by the omnivory motif increased significantly  with increasing connectance while the proportions of the other three motifs decreased (though not significantly).
        No proportions of motifs in the profile were significantly related to network size.

@@

    \begin{figure}[ht!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_vs_SC_lm.eps}
        \caption{The probabilities of persistence for a particular consumer (\textbf{A, B}) and the average probability of persistence across all consumers (\textbf{C-E}) varied with other elements of network structure as well as with motif participation.
        \textbf{A} Persistence of consumers decreased with increasing STL at all levels of disturbance.
        \textbf{B} Persistence of consumers increased with increasing in-degree when the probability of extinction for basal resources was low but decreased with increasing in-degree when the probability of extinction was high.
        \textbf{C,D} Mean persistence of consumers in a network decreased strongly with increasing probability of disturbance to basal resources and decreased slightly, but significantly, with increasing connectance. There was no significant relationship between mean persistence and network size.
        \textbf{E} Mean persistence of consumers decreased with increasing proportions of omnivory in the network's motif profile and increased with increasing proportions of other motifs. Persistence also decreased significantly with increasing probability of basal species extinction but there was no significant interaction; we therefore show trends only for $\pi$=0.1.}
        \label{fig:lm_CS}
    \end{figure}
